---
title: "DAE2"
author: "Dominic Sanchez"
date: "2026-04-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Read Displacement Test Data

```{r}
oil <- read.csv("CRUDEOIL.csv")
head(oil)
```


## Question 1: Histogram of Oil Recovery

```{r}
hist(oil$OILREC,
     main = "Histogram of Oil Recovery",
     breaks = 3,
     xlab = "Oil Recovery",
     ylab = "Frequency")
```


## Question 2: Scatterplot of Oil Recovery vs Dipping Angle

```{r}
plot(oil$DIPANGLE, oil$OILREC,
     main = "Scatterplot of Oil Recovery vs Dipping Angle",
     xlab = "Dipping Angle",
     ylab = "Oil Recovery")
```

Scatterplot shows a roughly linear trend.

## Question 3: Scatterplot of Oil Recovery vs Pressure

```{r}
plot(oil$PRESSURE, oil$OILREC,
     main = "Scatterplot of Pressure vs Oil Recovery",
     xlab = "Pressure",
     ylab = "Oil Recovery")
```

Scatterplot shows a moderately weak linear trend

## Question 4: Center the Independent Variables

```{r}
oil$angle.cen <- oil$DIPANGLE - mean(oil$DIPANGLE)
oil$pressure.cen <- oil$PRESSURE - mean(oil$PRESSURE)

head(oil)
```


## Question 5-12: Second-order model relating oil recovery to centered pressure and dipping angle

```{r}
oilreg.complete <- lm(OILREC ~ (angle.cen + pressure.cen)^2 + I(angle.cen^2) + I(pressure.cen^2), data = oil)
summary(oilreg.complete)
```

Collecting the sample size to calculate the degrees of freedom

```{r}
n <- nrow(oil)
n
```

## Question 13: Updated Regression Model

Removing the first variable (interaction term)

```{r}
oilreg.complete2 <- lm(OILREC ~ angle.cen + pressure.cen + I(angle.cen^2) + I(pressure.cen^2), data = oil)
summary(oilreg.complete2)
```

Second order effect of dip angle is still not significant, remove that next

```{r}
oilreg.complete3 <- lm(OILREC ~ angle.cen + pressure.cen + I(pressure.cen^2), data = oil)
summary(oilreg.complete3)
```

All of the coefficients are statistically significant, so the model is complete.


## Question 14: Residual plot

```{r}
plot(oilreg.complete3$fitted.values,
     oilreg.complete3$residuals,
     main = "Residuals vs Fitted Values",
     xlab = "Predicted Astringency",
     ylab = "Residuals")

abline(0,0)
```

**The residual plot shows no clear pattern and roughly constant spread, so this assumption is reasonable.**


## Question 15: Normal Quantile Plot

```{r}
qqnorm(oilreg.complete3$residuals)
qqline(oilreg.complete3$residuals)
```

**Ignoring the tail end points, the residuals mostly follow the line, so this assumption is reasonable.**



